Parsing Postfix log files

### 2008-11-22 doc/logparser-other-parsers-review.tex
Switched to new glossaries package. A couple of other XXX items have been done too.

Ten other parsers, starting with the only previously published research in
this area that the author is aware of, have been reviewed as part of the
background research for this project.  Each of the parsers was tested with
the \numberOFlogFILES{} test log files described in \sectionref{parser
efficiency}.  None of the parsers reviewed perform the kind of advanced
parsing and correlation required for this project; all except one (XXX IS
THIS ACCURATE\@?) are intended to perform a specific parsing and reporting
task, whereas \PLP{} is designed to be a generic parser, extracting data
and leaving generation of reports from the data to other programs.  Some
parsers save the data extracted to a data store, but the majority discard
all data once they have finished running, making historical analysis
impossible.  The parsers reviewed all produce a report of varying
complexity and detail, whereas this project's parser does not produce a
report at all; that responsibility is deferred to a separate program, to be
developed later.  The parser developed for this project is designed to
enable much more detailed log analysis by providing a stable platform for
subsequent programs to develop upon.

It is important to compare and contrast newly developed programs,
algorithms, and parsers against those already available, to accurately
judge what improvements, if any, are delivered by the newcomers.  There are
not that many previously developed Postfix log parsers --- indeed it was
quite difficult to find ten parsers to review for this project --- and the
functionality offered ranges from quite basic to much more mature,
depending on the needs of the creator.  The programs are not reviewed from
an independent viewpoint; the objective is to compare and contrast each
program with \parsername{}.  There are some important differences between
\parsername{} and the parsers reviewed here:

@@

    \item \parsername{} does not provide any reports based on the data it
        gathers.  This is by design: the parser is responsible for parsing
        log files and extracting data; further processing of that data is a
        separate concern, deferred to another program.  Using an \SQL{}
        database simplifies the process of generating such reports
        (discussed in \sectionref{database as API}); some simple examples
        are given in \sectionref{motivation}.

@@

XXX DESCRIBE THE \CSV{} FORMAT\@.

@@

There only appears to be one prior published paper about parsing Postfix
log files: \textit{Log Mail Analyzer: Architecture and Practical
Utilizations\/}~\cite{log-mail-analyser}.  The aim of \LMA{} is quite
different from this parser: it attempts to present correlated log data in a
form suitable for a systems administrator to search using the myriad of
standard Unix text processing utilities already available, producing both a
\CSV{} file and either a MySQL (\url{http://www.mysql.com/}) or Berkeley DB
(\url{http://www.oracle.com/database/berkeley-db/index.html}) database.
The decision to support both \CSV{} and Berkeley DB appears to have been a
serious limitation: \CSV{} is a very simple format where a record is stored
in a single line, with fields separated by a comma or other punctuation
symbol.  Problems with \CSV{} files include the need to escape separators
in the data stored, providing multiple values for a field (e.g.\ multiple
recipients), and adding new fields.  There is no mechanism to document the
fields or the separator, unlike \SQL{} databases where every database
includes a schema naming the fields and the type of data they store
(integer, text, timestamp, etc.).

Berkeley DB only supports storing simple \textbf{(key, value)} pairs,
unlike \SQL{} databases that store arbitrary tuples.  In \LMA{}'s main
table the key is an integer referred to by secondary tables, and the value
is a \CSV{} line containing all of the data for that row.  The
secondary by-sender, by-recipient, by-date, and by-\IP{} tables use the
sender/recipient/date/\IP{} as the key, and the value is a \CSV{} list of integers
referring to the main table.  This effectively re-implements
\SQL{} foreign keys, but without the functionality offered by even the most
basic of \SQL{} databases (joins, ordering, searches, etc.).  It also
requires custom code to search on some combination of the above, though the
authors of \LMA{} did provide some queries: IP-STORY, FROM-STORY,
DAILY-EMAIL and DAILY-REJECT\@.  Berkeley DB appears to be the least useful
of the three output formats: it does not provide the functionality of a
basic \SQL{} database, not can it be used with standard Unix text
processing tools.

\LMA{} stores the following date: time and date, client hostname and \IP{}
address, server \IP{} address, sender and recipient addresses, \SMTP{}
code, and size (for accepted mails only).  It omits XXX XXX\@.  Handling of
multiple recipients, \SMTP{} codes or remote servers\footnote{A single mail
may be sent to multiple remote servers if it was addressed to recipients in
different domains, or Postfix needs to try multiple servers for one or more
recipients.} is not explained in the paper; experimental observation shows
that multiple records are added when there are multiple recipients (sadly
the records are not associated or linked in any way), and presumably the
same approach is taken when there are multiple destination servers.

The schema used with the MySQL database is undocumented, but at least it is
possible to discover the schema with an \SQL{} database, unlike with
Berkeley DB\@; all \SQL{} databases embed the schema into the database and
provide commands for displaying it.  Berkeley DB does not embed a schema,
as there is neither requirement nor benefit; it only provides \textbf{(key,
value)} pairs, so any additional structuring of the data is imposed by the
application, thus it behoves the application to document this structure.

\LMA{} requires major changes to the parser code to parse new log lines or
extract additional data.  The code is structured as a long series of blocks
that each handle all lines matching a single regex, so parsing new log
lines requires modifying an existing regex or carefully inserting a new
block in the correct place; extracting extra data will require modifying
multiple blocks, regexes, or both.

\LMA{} does not appear to deal with any of the complications discussed in
\sectionref{complications}.  It does not differentiate between different
types of rejections, so it is not suitable for the purposes of this
project; the data about which restriction caused the rejection is
discarded, whereas the main goal of this project is to retain that data to
aid optimisation and evaluation of Postfix restrictions.  \LMA{} fails to
parse Postfix log files generated on Solaris hosts because the fields
automatically prepended to each log line differ from those added on Linux
hosts; log files from Solaris hosts (and possibly other operating systems)
thus require preprocessing before parsing by \LMA{}.  Once the
preprocessing has been performed on the \numberOFlogFILES{} test log files
\LMA{} parses the log files without complaint, although it produced 32
entries in its output file for every rejection in the input log file; it
also missed some 40\% of delivered mail.  Once these glaring deficiencies
were discovered the author did not waste any more time checking the
results.

\LMA{} does provide some simple reports: IP-STORY, FROM-STORY, DAILY-EMAIL
and DAILY-REJECT\@.  The first three extract \CSV{} lines for the specified
client \IP{} address, sender address, or date respectively.  DAILY-REJECT
initially failed with an error message (XXX ADD THE ERROR MESSAGE) from the
Perl interpreter; after making some corrections to the code it worked, and
extracted the \CSV{} lines for the specified day where the \SMTP{} code
signifies a rejection.  All of these reports are trivially simple to
produce from the \CSV{} file using the standard Unix tool \texttt{awk}; the
most complicated, DAILY-REJECT, is merely:

@@

    \item Some records have an \SMTP{} code of \textit{0}, though this is
        not a valid \SMTP{} code; presumably this is because \LMA{} did not
        parse a log line properly.  DAILY-REJECT considers this code to be
        a rejection, so the command above must too.

    \item In a small fraction of records the \SMTP{} code is
        \textit{deferred\/} (again, not a valid \SMTP{} code, and most
        likely because of mis-parsing by \LMA{}); it is treated as a
        rejection by DAILY-REJECT, so the equivalent \texttt{awk} command
        must do so, too.

@@

    \item \LMA{} does not escape or replace the separators used in the
        \CSV{} records when they are present in sender or recipient
        addresses, leading to some mangled records; these records are not
        present in the output from \texttt{awk}.  XXX WHY\@?  GIVE AN
        EXAMPLE RECORD\@.

    \item Where there are multiple records in the \CSV{} file, DAILY-REJECT
        extracts only one record, whereas the \texttt{awk} command above
        extracts all matching records.

@@

In summary, \LMA{} appears to be a proof of concept, written to demonstrate
the point of their paper (that having this information in an accessible
fashion is useful to systems administrators), rather than a program
designed to be useful in a production environment; no consideration was
given to extensibility when writing it.

% Literature review notes:
%
% Hard-coded parsing, requiring code changes to add more.  Attempts to
% correlate log lines, saves data to database for data mining purposes.
% Hard to extend/expand/understand.  Appears to only save: date and hour,
% DNS name and \IP{} address host, mail server \IP{} address, sender,
% receiver and e-mail status (sent, rejected).  Undocumented schema.
% Design decision to use \CSV{} as an intermediate format between the log
% file and the database seems to have been restrictive.  Appears to require
% a queueid but majority of log entries (e.g.\ rejections) lack a queueid.
% Supports whitelisting \IP{} addresses when parsing logs, but whitelisting
% when generating reports/data mining would be preferable.  Supporting
% Berkeley DB is probably limiting the software - an example is the
% difficulty in searching a pipe-delimited string, so they have
% re-implemented foreign keys with tables keyed by ip address etc.\
% pointing at the main table - this also will not scale well.  There does
% not appear to be any attempt to deal with the complications I have
% encountered: their parsing is not detailed enough to encounter them.  It
% does not run properly; does not create any output; throws up errors.

@@

Pflogsumm produces a report designed for troubleshooting rather than for
in-depth analysis.  It does not support saving any data or extracting more
data than is required for producing the report.  Both the parsing and
reporting are difficult to extend as it is a specialised tool, unlike the
easily extensible design of \parsername{}.  It does not correlate log lines
by queueid or \pid{}, and does not need to deal with the complications
encountered during this project.  Pflogsumm produces a useful report, and
successfully parsed the \numberOFlogFILES{} log files tested
with,\footnote{The results it reported were not verified in detail, but it
did not report any errors, and has a very good reputation amongst Postfix
users.} but it is not a suitable base for this project.

@@

Sawmill is a general purpose commercial product that parses 687 log file
formats (correct as of 2007/12/04) and produces reports.  Its data
extraction facilities are too limited to save sufficient data for the
purposes of this project,\footnote{The data extracted by Sawmill is
described later in this review, and the data extracted by \PLP{} is
described in \sectionref{connections table} and \sectionref{results
table}.} though it does extract three different sets of data for Postfix
(one is beta as of 2007/12/04), but they do not appear to be interlinked.
The documentation does not suggest that any attempt is made to correlate
log lines or deal with the difficulties documented in
\sectionref{complications}.  The source code is available in an obfuscated
form only (presumably for a fee), and the product is quite expensive,
requiring a license per report to be generated; in contrast \parsername{}
is free and the code is freely available.  The web interface allows
creation of dynamic reports based on any field, but because of the
associated cost the author has not experimented with it.  Presumably, as
the company charges per report the user wishes to generate, the data store
(if there is one) is deliberately inaccessible and undocumented to prevent
the user bypassing the program and generating their own reports.

XXX THE BLURB MENTIONS \SQL{} DB\@; CAN THAT BE ACCESSED SEPARATELY, IS THE
FORMAT DOCUMENTED, ETC\@?

\url{http://www.thesawmill.co.uk/formats/postfix.html} \newline{} Fields
extracted: from, to, server, UID, relay, status, number of recipients,
origin hostname, origin \IP{} and virus.  The fields \texttt{server},
\texttt{uid} and \texttt{virus} are not explained in the documentation:
\texttt{server} is probably the server the mail is delivered to, and
\texttt{uid} might be the uid of the user submitting mail locally.  Postfix
does not perform any form of virus checking (though it has many options for
cooperating with an external virus scanner), so the \texttt{virus} field is
a mystery.

\url{http://www.thesawmill.co.uk/formats/postfix_ii.html} \newline{} Fields
extracted: from, to, RBL, client hostname and client \IP{}\@.

\url{http://www.thesawmill.co.uk/formats/beta_postfix.html} \newline{} Fields
extracted: from, to, client hostname, client \IP{}, relay hostname, relay
\IP{}, status, response code, RBL and message id.

Even if the three data sets were linked together Sawmill would extract less
data than \parsername{}, and it does not appear to extract data about
rejections except when the rejection is caused by a \DNSBL{} check.

@@

When testing it took roughly 30 minutes to add a single log file to be
indexed (log files must be indexed before searching), with the downside
that the log file was copied into a spool directory, doubling the disk
space usage.  The most suitable test would be to index all the log files in
a particular directory, but after three hours, numerous futile attempts,
and reading all the available documentation the author admitted defeat.
Using the \CLI{} was more successful: invoking the command \newline{}
\tab{} \texttt{splunk find logs }\textit{log-directory\/}\newline{} added
40 of the \numberOFlogFILES{} log files to the queue for indexing.
Repeated attempts enqueued the same 40 log files, without explaining why
the others were excluded.\footnote{The log files appear to have been
indexed once only; presumably Splunk keeps track of the files it has
indexed and discards requests to index files for a second time.  This may
or may not be a useful feature for \parsername{}.} There did not appear to
be an option to order the log files to ensure they would be processed in
the order they were created, though this may not be necessary or beneficial
with Splunk.  Subsequently the author was successful in adding a single
file at a time using the \CLI{}; a simple loop to add all desired log files
was sufficient to index all files.  Splunk will periodically check all
those files for updates unless they are manually removed from its list;
this may or may not be useful behaviour.  Splunk did not appear to have any
difficulty in parsing the log files, once instructed to do so.
\parsername{} parses the logs it is instructed to parse, in the order
given; periodic parsing of logs is a task an administrator can easily
achieve with \texttt{cron(8)} and \texttt{logrotate(8)}.

@@

When parsing Postfix log files Splunk extracts some standard fields: to and
from addresses, HELO hostname, time and date, the host the log files were
generated on, and protocol (\SMTP{} or \ESMTP{}).  It parses the standard
syslog fields at the beginning of the log line, and extracts any
\texttt{key=<value>} pairs occurring after the standard syslog prologue;
these pairs are the fields listed above.  Searches can be based on the
extracted fields or the full text of the log line.  \parsername{} extracts
noticeably more data, though it does not make the full text of the line
available, and the full power of \SQL{} is available when searching,
allowing the user to search on arbitrarily complicated conditions.

Splunk is a generic tool, so it lacks any Postfix specific support over and
above extracting the \texttt{key=<value>} fields from a log line; most
importantly it makes no attempt to correlate log lines by queueid or
\pid{}, or to handle any of the myriad complications discussed in
\sectionref{complications}.  Some additional Postfix reports are supposedly
available at \url{http://www.splunkbase.com/}, but the author was unable to
find Postfix reports, or additional reports for any other log file types.

@@

AWStats will produce simple graphs for many different services, but
supporting many different services without special purpose code limits its
functionality.  The data it will extract from an \MTA{} log file is limited
in comparison to \parsername{}: \newline{} \tab{} time2, email, email\_r,
host, host\_r, method, url, code, and bytesd.\newline{} XXX WHAT FIELDS ARE
MISSING\@?  There does not seem to be an explanation of any of those fields
in the documentation (\parsername{} provides copious documentation).
AWStats coerces Postfix log files into Apache\footnote{The Apache web
server is the most popular HTTP server in use over the past 10 years; more
information is available at \url{http://httpd.apache.org/}.} format log
files, for analysis by AWStats' HTTP log file parser.  The converting
parser only deals with a small portion of the log lines generated by
Postfix, silently skipping those it cannot deal with, and does not
distinguish between different types of rejection; it would be a lot of work
to extend it to handle new log lines.  Although it does correlate log lines
by queueid (not by pid), it does not deal with any of the other
complications described in \sectionref{complications}.  AWStats supports
saving data but the format of the saved data is not documented, as far as
the author could tell.  It also supports reading compressed log files, but
that functionality was not tested.

@@

    \item A summary of the \SMTP{} codes used when rejecting delivery
        attempts.

@@

YAALA provides a mini-language based on \SQL{} that is used when generating
reports; sample reports can be seen
at~\url{http://www.yaala.org/samples.html}.  Example query for HTTP proxy
servers: \newline{} \tab{} \texttt{requests BY file WHERE host =\~{}
Google} \newline{} The mini-language is quite limited and cannot be used to
extract data for external use, merely to create reports.  Only data
selected by the query will be saved in the data store; other data will be
discarded, and removed from the data store if already present.

@@

Lire is a general purpose log parser supporting many different types of log
file.  The data extracted by its Postfix parser is shown in the quote
above; notably rejections are not mentioned or dealt with, and unlike
\parsername{} there is no facility to extend the parser.  It supports
multiple output formats for generated reports (text, HTML, PDF, and Excel
95) but the reports do not appear to be customisable; \parsername{} does
not produce any reports.  Lire produces a report with less detail than
Pflogsumm, and is considerable harder to configure.  Lire supports saving
extracted data for later report generation, but accessing this data from
another application is undocumented; given the source code it should be
possible, with enough time and effort, to understand the format.
\parsername{} uses an \SQL{} database to make accessing the extracted data
as easy as possible.

Like AWStats and Logrep, Lire attempts to correlate log lines by queueid,
but not by \pid{}, so the complete list of recipients for a mail should be
available; however its parser extracts only part of the available data and
makes no attempt to deal with the other complications described in
\sectionref{complications}.  When testing Lire on the \numberOFlogFILES{}
test log files it performed reasonably well, producing summaries of:

@@

    \item For mail sent: to addresses, \SMTP{} code, and delay.

@@

It also counts the number of log lines parsed and skipped.  Log lines are
correlated based on the queueid (referred to as sessionname [sic] within
Logrep), but not by \pid{}.  The parsing is error prone: empty fields are
saved when the log line does not match the regex, though it appears that
they will not overwrite existing data.  Most notably rejections are
completely ignored, making it unsuitable for the purposes of this project.
It does not try to address any of the complications in
\sectionref{complications} except for correlating by queueid.

Logrep does not come with any documentation, though some scant
documentation is available on its website (\parsername{} provides copious
documentation).  It requires a web browser to interact with it, so
automated log processing will be difficult, whereas automated processing is
a key part of \parsernames{} design.  Sadly all the author's attempts to
use Logrep failed, as it was unable to access the log files selected; this
appears to be a bug rather than operator error.  If it was caused by
operator error, the interface needs improvement as the (minimal)
instructions were followed as closely as possible, and multiple attempts
were made.  As parsing failed it was not possible to review the reports
Logrep can generate (available in HTML only), or to examine the
(undocumented) format in which it can save extracted data for subsequent
reuse.

>>>>

Ten other parsers, starting with the only previously published research in
this area that the author is aware of, have been reviewed as part of the
background research for this project.  Each of the parsers was tested with
the \numberOFlogFILES{} test log files described in \sectionref{parser
efficiency}.  None of the parsers reviewed perform the kind of advanced
parsing and correlation required for this project; all except one (XXX IS
THIS ACCURATE\@?) are intended to perform a specific parsing and reporting
task, whereas \gls{PLP} is designed to be a generic parser, extracting data
and leaving generation of reports from the data to other programs.  Some
parsers save the data extracted to a data store, but the majority discard
all data once they have finished running, making historical analysis
impossible.  The parsers reviewed all produce a report of varying
complexity and detail, whereas this project's parser does not produce a
report at all; that responsibility is deferred to a separate program, to be
developed later.  The parser developed for this project is designed to
enable much more detailed log file analysis by providing a stable platform
for subsequent programs to develop upon.

It is important to compare and contrast newly developed programs,
algorithms, and parsers against those already available, to accurately
judge what improvements, if any, are delivered by the newcomers.  There are
not that many previously developed Postfix log file parsers --- indeed it
was quite difficult to find ten parsers to review for this project --- and
the functionality offered ranges from quite basic to much more mature,
depending on the needs of the creator.  The programs are not reviewed from
an independent viewpoint; the objective is to compare and contrast each
program with \parsername{}.  There are some important differences between
\parsername{} and the parsers reviewed here:

@@

    \item \parsername{} does not provide any reports based on the data it
        gathers.  This is by design: the parser is responsible for parsing
        log files and extracting data; further processing of that data is a
        separate concern, deferred to another program.  Using an \gls{SQL}
        database simplifies the process of generating such reports
        (discussed in \sectionref{database as API}); some simple examples
        are given in \sectionref{motivation}.

@@

XXX DESCRIBE THE \gls{CSV} FORMAT\@.

@@

There only appears to be one prior published paper about parsing Postfix
log files: \textit{Log Mail Analyzer: Architecture and Practical
Utilizations\/}~\cite{log-mail-analyser}.  The aim of \gls{LMA} is quite
different from this parser: it attempts to present correlated data from log
files in a form suitable for a systems administrator to search using the
myriad of standard Unix text processing utilities already available,
producing both a \gls{CSV} file and either a MySQL
(\url{http://www.mysql.com/}) or Berkeley DB
(\url{http://www.oracle.com/database/berkeley-db/index.html}) database.
The decision to support both \gls{CSV} and Berkeley DB appears to have been
a serious limitation: \gls{CSV} is a very simple format where a record is
stored in a single line, with fields separated by a comma or other
punctuation symbol.  Problems with \gls{CSV} files include the need to
escape separators in the data stored, providing multiple values for a field
(e.g.\ multiple recipients), and adding new fields.  There is no mechanism
to document the fields or the separator, unlike \gls{SQL} databases where
every database includes a schema naming the fields and the type of data
they store (integer, text, timestamp, etc.).

Berkeley DB only supports storing simple \textbf{(key, value)} pairs,
unlike \gls{SQL} databases that store arbitrary tuples.  In \gls{LMA}'s
main table the key is an integer referred to by secondary tables, and the
value is a \gls{CSV} line containing all of the data for that row.  The
secondary by-sender, by-recipient, by-date, and by-\gls{IP} tables use the
sender/recipient/date/\gls{IP} as the key, and the value is a \gls{CSV}
list of integers referring to the main table.  This effectively
re-implements \gls{SQL} foreign keys, but without the functionality offered
by even the most basic of \gls{SQL} databases (joins, ordering, searches,
etc.).  It also requires custom code to search on some combination of the
above, though the authors of \gls{LMA} did provide some queries: IP-STORY,
FROM-STORY, DAILY-EMAIL and DAILY-REJECT\@.  Berkeley DB appears to be the
least useful of the three output formats: it does not provide the
functionality of a basic \gls{SQL} database, not can it be used with
standard Unix text processing tools.

\gls{LMA} stores the following date: time and date, client hostname and
\gls{IP} address, server \gls{IP} address, sender and recipient addresses,
\gls{SMTP} code, and size (for accepted mails only).  It omits XXX XXX\@.
Handling of multiple recipients, \gls{SMTP} codes or remote
servers\footnote{A single mail may be sent to multiple remote servers if it
was addressed to recipients in different domains, or Postfix needs to try
multiple servers for one or more recipients.} is not explained in the
paper; experimental observation shows that multiple records are added when
there are multiple recipients (sadly the records are not associated or
linked in any way), and presumably the same approach is taken when there
are multiple destination servers.

The schema used with the MySQL database is undocumented, but at least it is
possible to discover the schema with an \gls{SQL} database, unlike with
Berkeley DB\@; all \gls{SQL} databases embed the schema into the database
and provide commands for displaying it.  Berkeley DB does not embed a
schema, as there is neither requirement nor benefit; it only provides
\textbf{(key, value)} pairs, so any additional structuring of the data is
imposed by the application, thus it behoves the application to document
this structure.

\gls{LMA} requires major changes to the parser code to parse new log lines
or extract additional data.  The code is structured as a long series of
blocks that each handle all lines matching a single regex, so parsing new
log lines requires modifying an existing regex or carefully inserting a new
block in the correct place; extracting extra data will require modifying
multiple blocks, regexes, or both.

\gls{LMA} does not appear to deal with any of the complications discussed
in \sectionref{complications}.  It does not differentiate between different
types of rejections, so it is not suitable for the purposes of this
project; the data about which restriction caused the rejection is
discarded, whereas the main goal of this project is to retain that data to
aid optimisation and evaluation of Postfix restrictions.  \gls{LMA} fails
to parse Postfix log files generated on Solaris hosts because the fields
automatically prepended to each log line differ from those added on Linux
hosts; log files from Solaris hosts (and possibly other operating systems)
thus require preprocessing before parsing by \gls{LMA}.  Once the
preprocessing has been performed on the \numberOFlogFILES{} test log files
\gls{LMA} parses the log files without complaint, although it produced 32
entries in its output file for every rejection in the input log file; it
also missed some 40\% of delivered mail.  Once these glaring deficiencies
were discovered the author did not waste any more time checking the
results.

\gls{LMA} does provide some simple reports: IP-STORY, FROM-STORY,
DAILY-EMAIL and DAILY-REJECT\@.  The first three extract \gls{CSV} lines
for the specified client \gls{IP} address, sender address, or date
respectively.  DAILY-REJECT initially failed with an error message (XXX ADD
THE ERROR MESSAGE) from the Perl interpreter; after making some corrections
to the code it worked, and extracted the \gls{CSV} lines for the specified
day where the \gls{SMTP} code signifies a rejection.  All of these reports
are trivially simple to produce from the \gls{CSV} file using the standard
Unix tool \texttt{awk}\glsadd{awk}; the most complicated, DAILY-REJECT, is
merely:

@@

    \item Some records have an \gls{SMTP} code of \textit{0}, though this
        is not a valid \gls{SMTP} code; presumably this is because
        \gls{LMA} did not parse a log line properly.  DAILY-REJECT
        considers this code to be a rejection, so the command above must
        too.

    \item In a small fraction of records the \gls{SMTP} code is
        \textit{deferred\/} (again, not a valid \gls{SMTP} code, and most
        likely because of mis-parsing by \gls{LMA}); it is treated as a
        rejection by DAILY-REJECT, so the equivalent \texttt{awk} command
        must do so, too.

@@

    \item \gls{LMA} does not escape or replace the separators used in the
        \gls{CSV} records when they are present in sender or recipient
        addresses, leading to some mangled records; these records are not
        present in the output from \texttt{awk}.  XXX WHY\@?  GIVE AN
        EXAMPLE RECORD\@.

    \item Where there are multiple records in the \gls{CSV} file,
        DAILY-REJECT extracts only one record, whereas the \texttt{awk}
        command above extracts all matching records.

@@

In summary, \gls{LMA} appears to be a proof of concept, written to
demonstrate the point of their paper (that having this information in an
accessible fashion is useful to systems administrators), rather than a
program designed to be useful in a production environment; no consideration
was given to extensibility when writing it.

% Literature review notes:
%
% Hard-coded parsing, requiring code changes to add more.  Attempts to
% correlate log lines, saves data to database for data mining purposes.
% Hard to extend/expand/understand.  Appears to only save: date and hour,
% DNS name and \gls{IP} address host, mail server \gls{IP} address, sender,
% receiver and e-mail status (sent, rejected).  Undocumented schema.
% Design decision to use \gls{CSV} as an intermediate format between the
% log file and the database seems to have been restrictive.  Appears to
% require a queueid but majority of log lines (e.g.\ rejections) lack a
% queueid.  Supports whitelisting \gls{IP} addresses when parsing logs, but
% whitelisting when generating reports/data mining would be preferable.
% Supporting Berkeley DB is probably limiting the software - an example is
% the difficulty in searching a pipe-delimited string, so they have
% re-implemented foreign keys with tables keyed by ip address etc.\
% pointing at the main table - this also will not scale well.  There does
% not appear to be any attempt to deal with the complications I have
% encountered: their parsing is not detailed enough to encounter them.  It
% does not run properly; does not create any output; throws up errors.

@@

Pflogsumm produces a report designed for troubleshooting rather than for
in-depth analysis.  It does not support saving any data or extracting more
data than is required for producing the report.  Both the parsing and
reporting are difficult to extend as it is a specialised tool, unlike the
easily extensible design of \parsername{}.  It does not correlate log lines
by queueid or \gls{pid}, and does not need to deal with the complications
encountered during this project.  Pflogsumm produces a useful report, and
successfully parsed the \numberOFlogFILES{} log files tested
with,\footnote{The results it reported were not verified in detail, but it
did not report any errors, and has a very good reputation amongst Postfix
users.} but it is not a suitable base for this project.

@@

Sawmill is a general purpose commercial product that parses 687 log file
formats (correct as of 2007/12/04) and produces reports.  Its data
extraction facilities are too limited to save sufficient data for the
purposes of this project,\footnote{The data extracted by Sawmill is
described later in this review, and the data extracted by \gls{PLP} is
described in \sectionref{connections table} and \sectionref{results
table}.} though it does extract three different sets of data for Postfix
(one is beta as of 2007/12/04), but they do not appear to be interlinked.
The documentation does not suggest that any attempt is made to correlate
log lines or deal with the difficulties documented in
\sectionref{complications}.  The source code is available in an obfuscated
form only (presumably for a fee), and the product is quite expensive,
requiring a license per report to be generated; in contrast \parsername{}
is free and the code is freely available.  The web interface allows
creation of dynamic reports based on any field, but because of the
associated cost the author has not experimented with it.  Presumably, as
the company charges per report the user wishes to generate, the data store
(if there is one) is deliberately inaccessible and undocumented to prevent
the user bypassing the program and generating their own reports.

XXX THE BLURB MENTIONS \gls{SQL} DB\@; CAN THAT BE ACCESSED SEPARATELY, IS
THE FORMAT DOCUMENTED, ETC\@?

\url{http://www.thesawmill.co.uk/formats/postfix.html} \newline{} Fields
extracted: from, to, server, UID, relay, status, number of recipients,
origin hostname, origin \gls{IP} and virus.  The fields \texttt{server},
\texttt{uid} and \texttt{virus} are not explained in the documentation:
\texttt{server} is probably the server the mail is delivered to, and
\texttt{uid} might be the uid of the user submitting mail locally.  Postfix
does not perform any form of virus checking (though it has many options for
cooperating with an external virus scanner), so the \texttt{virus} field is
a mystery.

\url{http://www.thesawmill.co.uk/formats/postfix_ii.html} \newline{} Fields
extracted: from, to, RBL, client hostname and client \gls{IP}\@.

\url{http://www.thesawmill.co.uk/formats/beta_postfix.html} \newline{}
Fields extracted: from, to, client hostname, client \gls{IP}, relay
hostname, relay \gls{IP}, status, response code, RBL and message id.

Even if the three data sets were linked together Sawmill would extract less
data than \parsername{}, and it does not appear to extract data about
rejections except when the rejection is caused by a \gls{DNSBL} check.

@@

When testing it took roughly 30 minutes to add a single log file to be
indexed (log files must be indexed before searching), with the downside
that the log file was copied into a spool directory, doubling the disk
space usage.  The most suitable test would be to index all the log files in
a particular directory, but after three hours, numerous futile attempts,
and reading all the available documentation the author admitted defeat.
Using the \gls{CLI} was more successful: invoking the command \newline{}
\tab{} \texttt{splunk find logs }\textit{log-directory\/}\newline{} added
40 of the \numberOFlogFILES{} log files to the queue for indexing.
Repeated attempts enqueued the same 40 log files, without explaining why
the others were excluded.\footnote{The log files appear to have been
indexed once only; presumably Splunk keeps track of the files it has
indexed and discards requests to index files for a second time.  This may
or may not be a useful feature for \parsername{}.} There did not appear to
be an option to order the log files to ensure they would be processed in
the order they were created, though this may not be necessary or beneficial
with Splunk.  Subsequently the author was successful in adding a single
file at a time using the \gls{CLI}; a simple loop to add all desired log
files was sufficient to index all files.  Splunk will periodically check
all those files for updates unless they are manually removed from its list;
this may or may not be useful behaviour.  Splunk did not appear to have any
difficulty in parsing the log files, once instructed to do so.
\parsername{} parses the logs it is instructed to parse, in the order
given; periodic parsing of logs is a task an administrator can easily
achieve with \texttt{cron(8)} and \texttt{logrotate(8)}.

@@

When parsing Postfix log files Splunk extracts some standard fields: to and
from addresses, HELO hostname, time and date, the host the log files were
generated on, and protocol (\gls{SMTP} or \gls{ESMTP}).  It parses the
standard syslog\glsadd{syslog} fields at the beginning of the log line, and
extracts any \texttt{key=<value>} pairs occurring after the standard syslog
prologue; these pairs are the fields listed above.  Searches can be based
on the extracted fields or the full text of the log line.  \parsername{}
extracts noticeably more data, though it does not make the full text of the
line available, and the full power of \gls{SQL} is available when
searching, allowing the user to search on arbitrarily complicated
conditions.

Splunk is a generic tool, so it lacks any Postfix specific support over and
above extracting the \texttt{key=<value>} fields from a log line; most
importantly it makes no attempt to correlate log lines by queueid or
\gls{pid}, or to handle any of the myriad complications discussed in
\sectionref{complications}.  Some additional Postfix reports are supposedly
available at \url{http://www.splunkbase.com/}, but the author was unable to
find Postfix reports, or additional reports for any other log file types.

@@

AWStats will produce simple graphs for many different services, but
supporting many different services without special purpose code limits its
functionality.  The data it will extract from an \gls{MTA} log file is
limited in comparison to \parsername{}: \newline{} \tab{} time2, email,
email\_r, host, host\_r, method, url, code, and bytesd.\newline{} XXX WHAT
FIELDS ARE MISSING\@?  There does not seem to be an explanation of any of
those fields in the documentation (\parsername{} provides copious
documentation).  AWStats coerces Postfix log files into Apache\footnote{The
Apache web server is the most popular HTTP server in use over the past 10
years; more information is available at \url{http://httpd.apache.org/}.}
format log files, for analysis by AWStats' HTTP log file parser.  The
converting parser only deals with a small portion of the log lines
generated by Postfix, silently skipping those it cannot deal with, and does
not distinguish between different types of rejection; it would be a lot of
work to extend it to handle new log lines.  Although it does correlate log
lines by queueid (not by pid), it does not deal with any of the other
complications described in \sectionref{complications}.  AWStats supports
saving data but the format of the saved data is not documented, as far as
the author could tell.  It also supports reading compressed log files, but
that functionality was not tested.

@@

    \item A summary of the \gls{SMTP} codes used when rejecting delivery
        attempts.

@@

YAALA provides a mini-language based on \gls{SQL} that is used when
generating reports; sample reports can be seen
at~\url{http://www.yaala.org/samples.html}.  Example query for HTTP proxy
servers: \newline{} \tab{} \texttt{requests BY file WHERE host =\~{}
Google} \newline{} The mini-language is quite limited and cannot be used to
extract data for external use, merely to create reports.  Only data
selected by the query will be saved in the data store; other data will be
discarded, and removed from the data store if already present.

@@

Lire is a general purpose log file parser supporting many different types
of log file.  The data extracted by its Postfix parser is shown in the
quote above; notably rejections are not mentioned or dealt with, and unlike
\parsername{} there is no facility to extend the parser.  It supports
multiple output formats for generated reports (text, HTML, PDF, and Excel
95) but the reports do not appear to be customisable; \parsername{} does
not produce any reports.  Lire produces a report with less detail than
Pflogsumm, and is considerable harder to configure.  Lire supports saving
extracted data for later report generation, but accessing this data from
another application is undocumented; given the source code it should be
possible, with enough time and effort, to understand the format.
\parsername{} uses an \gls{SQL} database to make accessing the extracted
data as easy as possible.

Like AWStats and Logrep, Lire attempts to correlate log lines by queueid,
but not by \gls{pid}, so the complete list of recipients for a mail should
be available; however its parser extracts only part of the available data
and makes no attempt to deal with the other complications described in
\sectionref{complications}.  When testing Lire on the \numberOFlogFILES{}
test log files it performed reasonably well, producing summaries of:

@@

    \item For mail sent: to addresses, \gls{SMTP} code, and delay.

@@

It also counts the number of log lines parsed and skipped.  Log lines are
correlated based on the queueid (referred to as sessionname [sic] within
Logrep), but not by \gls{pid}.  The parsing is error prone: empty fields
are saved when the log line does not match the regex, though it appears
that they will not overwrite existing data.  Most notably rejections are
completely ignored, making it unsuitable for the purposes of this project.
It does not try to address any of the complications in
\sectionref{complications} except for correlating by queueid.

Logrep does not come with any documentation, though some scant
documentation is available on its website (\parsername{} provides copious
documentation).  It requires a web browser to interact with it, so
automated log file processing will be difficult, whereas enabling automated
processing is a key part of \parsernames{} design.  Sadly all the author's
attempts to use Logrep failed, as it was unable to access the log files
selected; this appears to be a bug rather than operator error.  If it was
caused by operator error, the interface needs improvement as the (minimal)
instructions were followed as closely as possible, and multiple attempts
were made.  As parsing failed it was not possible to review the reports
Logrep can generate (available in HTML only), or to examine the
(undocumented) format in which it can save extracted data for subsequent
reuse.
